---
title: "Class_09"
author: "Nancy Lin"
format: html
---

#The RCSB Protei Datat Base \>Q1: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy.


The webpage doesn't open up so we are skipping Q1-Q3.

>Q6

![![](images/1HSG%20(1).png){width="589"}](images/1HSG.png){width="443"}

> Picture 2 is with Asp25.
>
> Q4: Water molecules normally have 3 atoms. Why do we see just one atom per water molecule in this structure?

There is just one atom per water because the Hydrogen is too small to be seen. 

> Q5: There is a critical "conserved" water molecule in the binding site. Can you identify this water molecule? What residue number does this water molecule have

Yes, the residue number is 308. 


##Introduction to Bio3D in R

Bio3D is an R Package for structural Bioinformatics.To use it we need to call it up with the 'library()' function(just like any package).

```{r}
library(bio3d)
pdb <- read.pdb("1hsg")
pdb
```

> Q7: How many amino acid residues are there in this pdb object?

198 a.a

> Q8: Name one of the two non-protein residues?

HOH

> Q9: How many protein chains are in this structure?

There are two chains.

The atom records of the pbd files are stored in pdb\$stom

```{r}
head(pdb$atom)
```

> Q10. Which of the packages above is found only on BioConductor and not CRAN?

msa

> Q11. Which of the above packages is not found on BioConductor or CRAN?:

bio3d-view

> Q12. True or False? Functions from the devtools package can be used to install packages from GitHub and BitBucket?

True

#Comparative analysis ADK

Single PBD ID : 1AKE

First we get it's primary sequence:

```{r}
aa <- get.seq("1ake_a")
aa
```

> Q13. How many amino acids are in this sequence, i.e. how long is this sequence?

214a.a.

```{r}
b <- blast.pdb(aa)
```

```{r}
hits <- plot(b)

head(hits$pbd.id)
```

Use these ADK structures for analysis:

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','6S36_A','6RZE_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','3GMT_A','4PZL_A')
```

Download all these PBD files from the database...

```{r}
# Download releated PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

Align all the structures:

```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE,exefile="msa")
pdbs
```

Get_seq / Blast.pbd/ get.pbd / pbdaln

#Jump to PCA

```{r}
# Perform PCA
pc.xray <- pca(pdbs)
plot(pc.xray)
```

```{r}
# Calculate RMSD
rd <- rmsd(pdbs)

# Structure-based clustering
hc.rd <- hclust(dist(rd))
grps.rd <- cutree(hc.rd, k=3)

plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```

Visualization:

```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```

![](images/PC_1.PDB_animate-trajectory.mp4)

